### 2018-10-16 Manuscript/Manuscript.tex

%\usepackage{fontspec}
%\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[normalem]{ulem}
\usepackage[left]{lineno}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{color}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}
\usepackage{threeparttable}
\doublespacing

@@

XXX

@@

Outlines of LIPs through the Phanerozoic (Fig. \ref{fig:LIP_map}) were taken from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Fig. \ref{fig:LIP_map}; Table \ref{tab:LIPs}). The compilation of \citet{Ernst2017a} seeks to reconstruct the original areal extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs. These polygons encapsulate all of the rocks associated with a given LIP including dikes and sills (Fig. \ref{fig:LIP_map}). For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics. The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants--also creating the potential for them being over-estimates. However, this approach likely provide the best estimates available of original extent for ancient LIPs.  The extent of presently exposed volcanics associated with the LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.

After LIPs are emplaced they will progressively erode. In order to account for this decrease in area through with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function. They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time. We extend this approach to 19 basaltic LIPs for which we have estimates of the original areal extent of the province and the current areal extent of rocks associated with it (Fig. \ref{fig:LIP_preservation}). While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate approximate model for the progressive reduction in LIP area. The best-fit exponential function results in a LIP area half-life of 29~Myr. However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life, which yields a slightly higher best-fit half-life of 36~Myr (Fig. \ref{fig:LIP_preservation}). There is variability in the individual half-lives for LIPs such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps. In the analysis of LIP area through time, we implement decay scenarios informed by this analysis. Scenario `decay 1' uses the best-fit half-life of 36 Myr, while `decay 2' uses the slower decay with a half-life of $\sim$120~Myr. Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis. The LIP reconstruction used in this study includes pre-Phanerozoic LIPs. However, given the imposed exponential decay since emplacement, the inclusion of these LIPs do not significantly impact the calculated LIP areas throughout the Phanerozoic.

Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial by sediments of LIP volcanics that are co-located with a rift basin. There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}). For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}). To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting. `Burial 1' imposes instantaneous burial of 50$\%$ of the LIP area while `burial 2' imposes instantaneous burial of the entire LIP as an end-member scenario. The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios. Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects. In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g. the North American Mid-continent Rift).

@@

In the `decay 1' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig. \ref{fig:LIP_area}C). The first primary peak ca. 510~Ma is associated with the emplacement of the Kalkarindji LIP, the second primary peak ca. 380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third primary peak ca. 200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca. 30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca. 66~Ma Deccan LIP into the tropics (Fig. \ref{fig:LIP_area}A). When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca. 200~Ma peak is attenuated/removed due to the partial/complete removal of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP. However, after accounting for burial, a minor area of LIPs remain in the tropics from ca. 130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs. Using the longer decay half-life of 120~Myr (the `decay 2 + burial 2' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.

The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig. \ref{fig:LIP_area}C) and the ice extent record (Fig. \ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `decay 2 + burial 2' scenario; Fig. \ref{fig:LIP_correlation}). All other scenarios yield a weak negative correlation coefficient. Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordivician and Permo-Carboniferous glacial intervals.

To assess the statistical significance of the raw Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these simulations. For all scenarios except for the `decay 2 + burial 2' scenario, more than half of the randomly timed glacial intervals correlate better with the LIP area in the tropics (Fig. \ref{fig:LIP_correlation}). For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected.

@@

In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' Glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}. However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions. Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important (e.g. \citealp{Maher2014a}). As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}. Without active uplift, soil shielding from regolith development on low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability. In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate. This interpretation underlies the hypothesis that arc-continent collisons in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.

@@

We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500~Myr. Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until the build-up of sufficient pCO$_2$ to drive deglaciation. Nevertheless, the  hypothesis of tropical LIP area associated with the ca. 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth Glaciation is a major motivating driver behind conducting this analysis. How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic? Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator. This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the minor Cenozoic peak, and is significantly smaller than the other Phanerozoic peaks (Fig. \ref{fig:LIP_area}C). The ca. 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation. We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig. \ref{fig:LIP_area}C). Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, is not particularly notable, and therefore that the increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics is also not uniquely large, especially when considering the severity of the glaciation that followed it. This conclusion is consistent with the results of our Phanerozoic analysis, which suggests that LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.

\section*{CONCLUSIONS}

@@

@@

\begin{table}[h!]
\scriptsize
\caption{place holder}
\noindent\makebox[\textwidth]{
\begin{tabular}{lrlrrlrrl}
\toprule
                name &    age &                     age ref. &  original area Mkm2 &  present area Mkm2 &                                 present area ref. &  present/original &  half-life &   buried \\
\midrule
      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.1 &       no \\
                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al. (2006) &              0.31 &       17.8 &  partial \\
                NAIP &   62.0 &         Larsen et al. (2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al. (2006) &              0.27 &       32.8 &  partial \\
              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al. (2006) &              0.68 &      118.6 &       no \\
          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
          Madagascar &   90.0 &    Cucciniello et al. (2010) &                0.63 &               0.03 &                              Coffin et al. (2006) &              0.05 &       20.8 &       no \\
 Caribbean-Colombian &   94.0 &         Loewen et al. (2013) &                0.71 &               0.13 &                              Coffin et al. (2006) &              0.18 &       38.0 &       no \\
               HALIP &   95.0 &      Kingsbury et al. (2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.5 &       no \\
              EQUAMP &  131.0 &       Hollanda et al. (2016) &                0.66 &               0.01 &                            Hollanda et al. (2016) &              0.01 &       19.7 &       no \\
               Comei &  132.0 &            Zhu et al. (2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
             Bunbury &  132.0 &            Zhu et al. (2009) &                0.03 &               0.00 &                              Thorne et al. (2014) &              0.05 &       30.5 &       no \\
     Parana-Etendeka &  135.0 &      Florisbal et al. (2014) &                3.12 &               0.40 &  Coffin et al. (2016), Teixeira and Albrez (2017) &              0.13 &       45.9 &  partial \\
                Trap &  140.0 &          Ernst et al. (2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
 NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
               Karoo &  183.0 &        Burgess et al. (2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       42.3 &       no \\
              Ferrar &  183.0 &        Burgess et al. (2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
                CAMP &  201.0 &      Blackburn et al. (2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.6 &  partial \\
             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al. (2006) &              0.14 &       88.8 &       no \\
            Emeishan &  259.0 &           Zhou et al. (2002) &                0.71 &               0.06 &                              Coffin et al. (2006) &              0.09 &       74.6 &       no \\
    Panjal-Qiangtang &  283.0 &           Zhai et al. (2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
               Tarim &  290.0 &             Xu et al. (2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Magdalen &  360.0 &         Murphy et al. (1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
              Vilyui &  374.0 &          Renne et al. (2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
        Kola-Dnieper &  380.0 &    Arzamastsev and Wu (2014) &                5.90 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Suordakh &  450.0 &       Khudoley et al. (2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
         Kalkarindji &  511.0 &        Jourdan et al. (2014) &                3.54 &               0.17 &                              Thorne et al. (2014) &              0.05 &      118.2 &       no \\
            Franklin &  720.0 &       Denyszyn et al. (2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      127.6 &       no \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
place holder.
\end{tablenotes}
\label{tab:LIPs}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.65\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
	\caption{LIP preservation through time. The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs. An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement, which yields a half-life of $\sim$36~Myr.}
	\label{fig:LIP_preservation}
\end{center}
\end{figure}

>>>>

%\usepackage{fontspec}
%\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[normalem]{ulem}
\usepackage[left]{lineno}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{color}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}
\usepackage{threeparttable}
\doublespacing

@@

One of the hypothesized effects of the emplacement of large igneous provinces (LIPs) is planetary cooling on million-year timescales associated with enhanced silicate weathering. A systematic analysis of LIP area through the Phanerozoic reveals that there is not a significant correlation between total LIP area, nor LIP area in the tropics, and the extent of continental ice sheets. The largest peaks in tropical LIP area are at times of non-glacial climate. Changes in planetary weatherability associated with LIPs are therefore not the fundamental control on whether Earth is in a glacial or non-glacial climate although they could provide a secondary modulating effect in conjunction with other processes.

@@

Outlines of LIPs through the Phanerozoic (Fig. \ref{fig:LIP_map}) were taken from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Fig. \ref{fig:LIP_map}; Table \ref{tab:LIPs}). The compilation of \citet{Ernst2017a} seeks to reconstruct the original areal extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs. These polygons encapsulate all of the rocks associated with a given LIP including dikes and sills (Fig. \ref{fig:LIP_map}). For some LIPs, this approach may lead to an overestimate of original areal extent given that subsurface intrusions could extend over a broader area than the surface volcanics. The polygons also assume complete areal coverage by LIPs as they connect wide-spread remnants--also creating the potential for them being over-estimates. However, this approach likely provide the best estimates available of original extent for ancient LIPs.  The extents of presently-exposed volcanics associated with LIPs were taken from a number of resources including the PLATES compilation (Table \ref{tab:LIPs}; \citealp{Coffin2006a}) and more recent compilation efforts associated with the Large Igneous Provinces Commission.

After LIPs are emplaced, they will progressively erode. In order to account for this decrease in area through with time, \citet{Godderis2017a} took the approach of fitting estimates of the original surface extent and estimates of the current extent for 5 LIPs with an exponential decay function. They used the resulting exponential decay constants to develop a first-order model to estimate changing LIP area through time. We extend this approach to 19 basaltic LIPs for which we have estimates of the original areal extent of the province and the current areal extent of rocks associated with it (Fig. \ref{fig:LIP_preservation}). While there are significant uncertainties associated with the area estimates, this compilation suggests that an exponential decay function is an appropriate approximate model for the progressive reduction in LIP area. The best-fit exponential function results in a LIP area half-life of 29~Myr. However, since we explicitly account for LIP burial separately in our models (see below), we prefer to exclude the 6 of 19 LIPs which have been partially or completely buried in our estimation of a representative LIP area half-life, which yields a slightly higher best-fit half-life of 36~Myr (Fig. \ref{fig:LIP_preservation}). There is variability in the half-lives for LIPs, if they are each individual fit with an exponential decay function, such that they range from shorter half-lives of $\sim$20~Myr up to $\sim$120~Myr for the Kalkarindji and Deccan Traps. In the analysis of LIP area through time, we implement decay scenarios informed by this analysis. Scenario `decay 1' uses the best-fit half-life of 36 Myr, while `decay 2' uses the slower decay with a half-life of $\sim$120~Myr. Given that the post-emplacement weathering and erosional history of each LIP should be dependent on the tectonic and climatic setting that each LIP experiences during and after emplacement, this approach is simplistic but provides a framework for analysis. The LIP reconstruction used in this study includes pre-Phanerozoic LIPs. However, given the imposed exponential decay since emplacement, the inclusion of these LIPs does not significantly impact the calculated LIP areas throughout the Phanerozoic Eon which is the focus of this analysis.

Of the tectonic factors that could alter exposed LIP area, the most consequential is near-immediate burial by sediments of LIP volcanics that are co-located with a rift basin. There are numerous examples in the record where there is partial or complete burial of a LIP associated with rifting and thermal subsidence (Table \ref{tab:LIPs}). For example, the Afar LIP is both associated with the Ethiopian Traps which form plateau flood basalts, and successful rifting in the region of the Red Sea that has resulted in burial (Fig \ref{fig:LIP_map}). To account for the rapid decrease in exposed surface area that would result from burial by sediments in a rift basin, we impose two different burial scenarios for LIPs that are co-located with rifting. `Burial 1' imposes instantaneous burial of 50$\%$ of the LIP area while `burial 2' imposes instantaneous burial of the entire LIP as an end-member scenario. The LIP area analysis uses all of the available combinations of these distinct decay and burial scenarios. Our LIP reconstruction differs from that of \citet{Johansson2018a} in several respects. In contrast to the decay and decay+burial scenarios implemented on estimates of original LIP extent in this study, \citet{Johansson2018a} uses a static extent for each LIP throughout the reconstruction with some of the polygons corresponding to the present-day surface extent and some representing the original extent that includes currently buried portions of the LIP (e.g. the North American Midcontinent Rift).

@@

In the `decay 1' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig. \ref{fig:LIP_area}C). The first primary peak ca. 510~Ma is associated with the emplacement of the Kalkarindji LIP, the second primary peak ca. 380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third primary peak ca. 200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca. 30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca. 66~Ma Deccan LIP into the tropics (Fig. \ref{fig:LIP_area}A). When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca. 200~Ma peak is attenuated/removed due to the partial/complete burial of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP. However, after accounting for burial, a minor area of LIPs remain in the tropics from ca. 130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs. Using the longer decay half-life of 120~Myr (the `decay 2 + burial 2' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.

The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig. \ref{fig:LIP_area}C) and the ice extent record (Fig. \ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `decay 2 + burial 2' scenario; Fig. \ref{fig:LIP_correlation}). All other scenarios yield a near zero or weak negative correlation coefficient. Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordovician and Permo-Carboniferous glacial intervals.

To assess the statistical significance of the raw Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations.  For the `decay 2 + burial 2' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected (Fig. \ref{fig:LIP_correlation}). Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.

@@

In the original models that proposed the `Fire and Ice' hypothesis as an explanation for the onset of the Sturtian `Snowball Earth' glaciation, chemical weathering was modeled as an Arrhenius relationship where it was solely a function of temperature and run-off \citep{Donnadieu2004a}. However, such an approach neglects the effects that can result through soil shielding and regolith development in low-relief regions. Recent progress on understanding the relationships between landscapes, topography, and chemical weathering reveals that these effects are quite important (e.g. \citealp{Maher2014a}). As a result, more recent modeling of chemical weathering that has incorporated such processes highlights the importance of high-relief regions relative to low-relief ones \citep{Godderis2017b}. Without active uplift, soil shielding from regolith development on low-relief LIPs could significantly decrease the local weatherability of a LIP and mute its impact on global weatherability. In contrast, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography, particularly in tropical regions, may exert a strong control on global weatherability and long-term climate. This interpretation underlies the hypothesis that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a} and the Cenozoic \citep{Jagoutz2016a} played a significant role in transitions into glacial climate states at those times.

@@

We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$500~Myr. Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient pCO$_2$ builds up in the absence of silicate weathering to drive deglaciation. Nevertheless, the hypothesis of tropical LIP area associated with the ca. 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth Glaciation is a major motivating driver behind conducting this analysis. How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic? Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator. This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the minor Cenozoic peak, and is significantly smaller than the other Phanerozoic peaks (Fig. \ref{fig:LIP_area}C). The ca. 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation. We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig. \ref{fig:LIP_area}C). Together, these results indicate that the tropical area of the Franklin LIP, when compared that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large are in the tropics. Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth. One such process could have been an increase in global weatherability associated with the accretion of arc terranes within terranes of the present-day Arabian-Nubian shield. An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor on top of other such processes. The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling. This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation.   injection is also not uniquely large, especially when considering the severity of the glaciation that followed it. This conclusion is consistent with the results of our Phanerozoic analysis, which suggests that LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.

@@

\clearpage
\newpage

@@

\begin{table}[h!]
\scriptsize
\caption{Phanerozoic (and the Franklin) large igneous provinces.}
\resizebox{\textwidth}{!}{
\begin{tabular}{lc>{\raggedright}p{3.5cm}cc>{\raggedright}p{3.5cm}ccl}
\toprule
name & age & age ref. & original & present & present area ref. & present/ & half-life & buried? \\
 & (Ma) & & area & area & & original & (Myr) & \\
 & & & (Mkm$^{2}$) & (Mkm$^{2}$) & & & & \\
\midrule
Columbia River & 16 & Kasbohm and Schoene (2018) & 0.68 & 0.38 & Buchan and Ernst (2004) & 0.56 & 19.2 & no \\
Afar & 30 & Courtillot and Renne (2003) & 2.05 & 0.63 & Coffin et al. (2006) & 0.31 & 17.7 & partial \\
NAIP & 62 & Larsen et al. (2016) & 1.07 & 0.29 & Buchan and Ernst (2004), Coffin et al. (2006) & 0.27 & 33.0 & partial \\
Deccan & 66 & Schoene (2015) & 0.83 & 0.56 & Coffin et al. (2006) & 0.68 & 116.6 & no \\
Seychelles & 66 & Schoene (2015) & 0.46 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
Madagascar & 90 & Cucciniello et al. (2010) & 0.63 & 0.03 & Coffin et al. (2006) & 0.05 & 20.8 & no \\
Caribbean-Colombian & 94 & Loewen et al. (2013) & 0.71 & 0.13 & Coffin et al. (2006) & 0.18 & 37.6 & no \\
HALIP & 95 & Kingsbury et al. (2018) & 3.60 & 0.15 & Moosdorf and Hartmann (2015) & 0.04 & 20.7 & no \\
EQUAMP & 131 & Hollanda et al. (2016) & 0.66 & 0.01 & Hollanda et al. (2016) & 0.01 & 20.5 & no \\
Comei & 132 & Zhu et al. (2009) & 0.11 & - & - & - & - & no \\
Bunbury & 132 & Zhu et al. (2009) & 0.03 & 0.00 & Thorne et al. (2014) & 0.05 & 30.9 & no \\
Parana-Etendeka & 135 & Florisbal et al. (2014) & 3.12 & 0.40 & Coffin et al. (2016), Teixeira and Albrez (2017) & 0.13 & 45.7 & partial \\
Trap & 140 & Ernst et al. (2001) & 0.03 & 0.00 & Ernst and Buchan (2001) & 0.00 & 0.0 & no \\
NW Australia Margin & 160 & COME BACK TO THIS & 0.62 & 0.00 & Coffin et al. (2006) & 0.00 & 0.0 & yes \\
Karoo & 183 & Burgess et al. (2015) & 3.21 & 0.15 & de Kock compilation & 0.05 & 41.3 & no \\
Ferrar & 183 & Burgess et al. (2015) & 0.18 & - & - & - & - & no \\
CAMP & 201 & Blackburn et al. (2013) & 11.46 & 0.23 & Marzoli and Parisio compilation & 0.02 & 35.7 & partial \\
Siberia & 252 & Burgess and Bowring (2015) & 3.46 & 0.47 & Coffin et al. (2006) & 0.14 & 87.5 & no \\
Emeishan & 259 & Zhou et al. (2002) & 0.71 & 0.06 & Coffin et al. (2006) & 0.09 & 72.9 & no \\
Panjal-Qiangtang & 283 & Zhai et al. (2013) & 0.11 & - & - & - & - & no \\
Tarim & 290 & Xu et al. (2014) & 0.35 & - & - & - & - & no \\
Magdalen & 360 & Murphy et al. (1999) & 0.42 & - & - & - & - & no \\
Vilyui & 374 & Renne et al. (2010) & 1.14 & - & - & - & - & no \\
Kola-Dnieper & 380 & Arzamastsev and Wu (2014) & 5.90 & - & - & - & - & no \\
Suordakh & 450 & Khudoley et al. (2013) & 0.02 & - & - & - & - & no \\
Kalkarindji & 511 & Jourdan et al. (2014) & 3.54 & 0.17 & Thorne et al. (2014) & 0.05 & 116.3 & no \\
Franklin & 720 & Denyszyn et al. (2009) & 2.62 & 0.04 & Buchan and Ernst (2004) & 0.02 & 121.8 & no \\
\bottomrule
\end{tabular}
}
\label{tab:LIPs}
\end{table}

\clearpage
\newpage

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.6\textwidth]{Manuscript/Figures/LIP_Preservation.pdf}
	\caption{LIP preservation through time. The ratio of estimates of the present-day area to that of the original area are shown for 19 basaltic LIPs. An exponential fit is made to the 13 basaltic LIPs that have not been buried after emplacement, which yields a half-life of $\sim$36~Myr.}
	\label{fig:LIP_preservation}
\end{center}
\end{figure}
